Analysis Script

### 2025-05-02 code/analysis-code/statisitical-analysis.qmd
Edits

This code chunk merges the results of the best models into a data frame.
```{r}
# Reformat these data frames for joining

base_lasso_metrics <- all_base_lasso_metrics %>% 
  mutate(model = "LASSO (Baseline)")

@@

```{r}
model_table <- model_metrics %>%
  select(model, rmse_Train, mae_Train, rsq_Train, rmse_Test, mae_Test, rsq_Test) %>%
  gt() %>%
  tab_header(
    title = "Model Performance Metrics",
    subtitle = "Train and Test Results | Best Performance Highlighted in Green"
  ) %>%
  fmt_number(columns = everything(), decimals = 3) %>%
  cols_label(
    model = "Model",
    rmse_Train = "RMSE",
    mae_Train = "MAE",
    rsq_Train = "R²",
    rmse_Test = "RMSE",
    mae_Test = "MAE",
    rsq_Test = "R²"
  ) %>%
  tab_spanner(
    label = "Train Metrics",
    columns = c(rmse_Train, mae_Train, rsq_Train)
  ) %>%
  tab_spanner(
    label = "Test Metrics",
    columns = c(rmse_Test, mae_Test, rsq_Test)
  ) %>%
  tab_style(
    style = cell_fill(color = "lightgreen"),
    locations = cells_body(
      columns = rmse_Test,
      rows = rmse_Test == min(rmse_Test, na.rm = TRUE)
    )
  ) %>%
  tab_style(
    style = cell_fill(color = "lightgreen"),
    locations = cells_body(
      columns = mae_Test,
      rows = mae_Test == min(mae_Test, na.rm = TRUE)
    )
  ) %>%
  tab_style(
    style = cell_fill(color = "lightgreen"),
    locations = cells_body(
      columns = rsq_Test,
      rows = rsq_Test == max(rsq_Test, na.rm = TRUE)
    )
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(everything())
  ) %>%
  tab_options(
    table.width = pct(100),
    table.align = "left",
    column_labels.padding = px(10),
    data_row.padding = px(6)
  )

@@

This code chunk creates a short summary table of the results from the model_metrics data frame.
```{r}
# Create gt table
model_table <- model_metrics %>%
  gt() %>%
  tab_header(
    title = "Model Performance Metrics on Data",
    subtitle = "Best values are highlighted per metric"
  ) %>%
  fmt_number(columns = c(rmse, mae), decimals = 1) %>%
  fmt_number(columns = rsq, decimals = 3) %>%
  cols_label(
    model = "Model",
    rmse = "RMSE",
    rsq = "R²",
    mae = "MAE"
  ) %>%

  # Highlight lowest RMSE
  tab_style(
    style = cell_fill(color = "lightgreen"),
    locations = cells_body(
      columns = rmse,
      rows = rmse == min(rmse, na.rm = TRUE)
    )
  ) %>%

  # Highlight lowest MAE
  tab_style(
    style = cell_fill(color = "lightgreen"),
    locations = cells_body(
      columns = mae,
      rows = mae == min(mae, na.rm = TRUE)
    )
  ) %>%

  # Highlight highest R²
  tab_style(
    style = cell_fill(color = "lightgreen"),
    locations = cells_body(
      columns = rsq,
      rows = rsq == max(rsq, na.rm = TRUE)
    )
  ) %>%

  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(everything())
  ) %>%
  tab_options(
    table.width = pct(100),
    table.align = "left",
    column_labels.padding = px(10),
    data_row.padding = px(6)
  )

>>>>

This code chunk merges the results of the best models into a data frame.
```{r}
# Reformat these data frames for joining
base_lasso_metrics <- all_base_lasso_metrics %>% 
  mutate(model = "LASSO (Baseline)")

@@

This code chunk creates a short summary table of the results from the model_metrics data frame.

```{r}
# Create table and also reorder data frame columns
model_table <- model_metrics %>%
  select(model, rmse_Train, mae_Train, rsq_Train, rmse_Test, mae_Test, rsq_Test) %>%
  gt() %>%
  tab_header(
    title = "Model Performance Metrics",
    subtitle = "Train and Test Results | Best Performance Highlighted in Green"
  ) %>%
  fmt_number(columns = everything(), decimals = 3) %>%
  cols_label(
    model = "Model",
    rmse_Train = "RMSE",
    mae_Train = "MAE",
    rsq_Train = "R²",
    rmse_Test = "RMSE",
    mae_Test = "MAE",
    rsq_Test = "R²"
  ) %>%
  tab_spanner(
    label = "Train Metrics",
    columns = c(rmse_Train, mae_Train, rsq_Train)
  ) %>%
  tab_spanner(
    label = "Test Metrics",
    columns = c(rmse_Test, mae_Test, rsq_Test)
  ) %>%
  tab_style(
    style = cell_fill(color = "lightgreen"),
    locations = cells_body(
      columns = rmse_Test,
      rows = rmse_Test == min(rmse_Test, na.rm = TRUE)
    )
  ) %>%
  tab_style(
    style = cell_fill(color = "lightgreen"),
    locations = cells_body(
      columns = mae_Test,
      rows = mae_Test == min(mae_Test, na.rm = TRUE)
    )
  ) %>%
  tab_style(
    style = cell_fill(color = "lightgreen"),
    locations = cells_body(
      columns = rsq_Test,
      rows = rsq_Test == max(rsq_Test, na.rm = TRUE)
    )
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(everything())
  ) %>%
  tab_options(
    table.width = pct(100),
    table.align = "left",
    column_labels.padding = px(10),
    data_row.padding = px(6)
  )

@@
